Preprocessing

### 2023-08-29 analysis/preprocessing.Rmd
Update preprocessing.Rmd

```{r}
d_journals <- d_journals %>%
  select(timestamps = Timestamp,
         coder_id = `Coder's Initials`,
         journal = `Journal Name`,
         journal_empirical = `Does the journal publish empirical research?`,
         anyPPC = `Check the article types page and check for web comments — does the journal offer any type of PPC?`,
         A_ppc_name = `A. Enter the name of the PPC type offered by the journal:`,
         B_ppc_name = `B. Enter the name of the PPC type offered by the journal:`,
         C_ppc_name = `C. Enter the name of the PPC type offered by the journal:`,
         A_ppc_description = `A. Enter the verbatim description of this type of PPC provided by the journal`,
         B_ppc_description = `B. Enter the verbatim description of this type of PPC provided by the journal`,
         C_ppc_description = `C. Enter the verbatim description of this type of PPC provided by the journal`,
         A_ppc_length = `A. Are there any length limits for this type of PPC?`,
         B_ppc_length = `B. Are there any length limits for this type of PPC?`,
         C_ppc_length = `C. Are there any length limits for this type of PPC?`,
         A_ppc_time = `A. Are there any time limits for submission of this type of PPC?`,
         B_ppc_time = `B. Are there any time limits for submission of this type of PPC?`,
         C_ppc_time = `C. Are there any time limits for submission of this type of PPC?`,
         A_ppc_ref = `A. Are there any reference limits for this type of PPC?`,
         B_ppc_ref = `B. Are there any reference limits for this type of PPC?`,
         C_ppc_ref = `C. Are there any reference limits for this type of PPC?`,
         A_ppc_review = `A. Is this type of PPC sent for independent external peer review?`,
         B_ppc_review = `B. Is this type of PPC sent for independent external peer review?`,
         C_ppc_review = `C. Is this type of PPC sent for independent external peer review?`,
         A_ppc_note = `A. Anything additional, unusual, or interesting to note?`,
         B_ppc_note = `B. Anything additional, unusual, or interesting to note?`,
         C_ppc_note = `C. Anything additional, unusual, or interesting to note?`,
         general_note = `Anything additional, unusual, or interesting to note?`,
         -starts_with("D. "),
         -ends_with("Are there any other types of PPC in this journal? (Remember to check for web comments!)`"))

d_articles <- d_articles %>%
  select(
    article_id = `article id`,
    coder_id_primary = `primary coder initials`,
    coder_id_secondary = `secondary coder initials`,
    exclusion_primary = `1_exclusion`,
    exclusion_reason_primary = `1_exclusionReason`,
    ppc_linked_primary = `1_Is article linked to PPC`,
    notes_primary = `1_Additional notes`,
    exclusion_secondary = `2_exclusion`,
    exclusion_reason_secondary = `2_exclusionReason`,
    ppc_linked_secondary = `2_Is article linked to PPC`,
    notes_secondary = `2_Additional notes`
  )
```

@@

Some journal names were entered in the extraction form with incorrect formatting. Correct this below before running the test.

```{r}
d_journals <- d_journals %>% 
  mutate(journal = fct_recode(journal,
                              "MEMORY & COGNITION" = "Memory & Cognition","JOURNAL OF COUPLE & RELATIONSHIP THERAPY-INNOVATIONS IN CLINICAL AND EDUCATIONAL INTERVENTIONS" = "Journal of Couple & Relationship Therapy"))
```

@@

# Attach sample information to extracted data

## Identify sample for journals

The `d_journals` dataset does not identify whether journals were in the prominent sample, random sample, or both. So we need to get the sample ID retroactively using the files containing all journals included in the sample.

Create sample ID column(s) to indicate whether in prominent, random, or both samples

```{r}
d_journals <- d_journals %>%
  mutate(
    sample_journal_prominent = journal %in% journal_list_prominent$Journal,
    sample_journal_random = journal %in% journal_list_random$Journal,
    sample_journal_both = sample_journal_prominent & sample_journal_random,
    sample_journal_error = !sample_journal_prominent & !sample_journal_random
  )

#Creates sampleID column denoting; prominent, random, or both to replace the 4 specified columns above
d_journals <- d_journals %>% 
  mutate(sample_journal_prominent = ifelse(sample_journal_prominent, "prominent", "random")) %>%
  mutate(sample_journal_both = ifelse(sample_journal_both, "both", "na")) %>%
  mutate(sampleID = case_when(
    sample_journal_both == "both" ~ "both",
    sample_journal_both == "na" ~ sample_journal_prominent,
    TRUE ~ sample_journal_both
  )) %>%
  select(-c(sample_journal_prominent, sample_journal_random, sample_journal_both, sample_journal_error))
```

# Resolve coding disagreements in articles data

@@

```{r}
# create exclusion_final column
d_articles <- d_articles %>%
  mutate(exclusion_final = case_when(
    sampleID == 'random' & article_id == '37' ~ exclusion_primary,
    sampleID == 'random' & article_id == '41' ~ exclusion_primary,
    sampleID == 'random' & article_id == '68' ~ exclusion_secondary,
    sampleID == 'random' & article_id == '95' ~ exclusion_secondary,
    sampleID == 'random' & article_id == '103' ~ exclusion_primary,
    sampleID == 'random' & article_id == '120' ~ exclusion_primary,
    sampleID == 'prominent' & article_id == '30' ~ exclusion_secondary,
    sampleID == 'prominent' & article_id == '49' ~ exclusion_primary,
    sampleID == 'prominent' & article_id == '66' ~ exclusion_primary,
    sampleID == 'prominent' & article_id == '98' ~ exclusion_primary,
    TRUE ~ exclusion_primary
  ))

# create exclusion_reason_final column
d_articles <- d_articles %>%
  mutate(exclusion_reason_final = case_when(
    sampleID == 'random' & article_id == '37' ~ exclusion_reason_primary,
    sampleID == 'random' & article_id == '41' ~ exclusion_reason_primary,
    sampleID == 'random' & article_id == '68' ~ exclusion_reason_secondary,
    sampleID == 'random' & article_id == '95' ~ exclusion_reason_secondary,
    sampleID == 'random' & article_id == '103' ~ exclusion_reason_primary,
    sampleID == 'random' & article_id == '120' ~ exclusion_reason_primary,
    sampleID == 'prominent' & article_id == '30' ~ exclusion_reason_secondary,
    sampleID == 'prominent' & article_id == '49' ~ exclusion_reason_primary,
    sampleID == 'prominent' & article_id == '66' ~ exclusion_reason_primary,
    sampleID == 'prominent' & article_id == '98' ~ exclusion_reason_primary,
    TRUE ~ exclusion_reason_primary
  ))
```

@@

```{r}
# we've already dealt with cases where the decision to exclude differs, so we're not looking only at cases where the decision to exclude matches, but the reason for exclusion differs
d_articles %>%
  filter(exclusion_primary == exclusion_secondary, 
         exclusion_reason_primary != exclusion_reason_secondary) %>%
  select(article_id, sampleID, exclusion_reason_primary, exclusion_reason_secondary)
```
We will resolve these by applying the exclusion criteria hierarchically. In other words, we use the highest level of exclusion criteria assigned by one of the two coders.
1. Access
2. English
3. Retracted 
4. Non-empirical

```{r}
# update exclusion_reason_final column when primary and secondary coder disagree on reason to exclude.
d_articles <- d_articles %>%
  mutate(exclusion_reason_final = case_when(
    sampleID == 'random' & article_id == '14' ~ exclusion_reason_primary,
    sampleID == 'random' & article_id == '35' ~ exclusion_reason_primary,
    sampleID == 'random' & article_id == '54' ~ exclusion_reason_secondary,
    sampleID == 'random' & article_id == '73' ~ exclusion_reason_primary,
    sampleID == 'random' & article_id == '78' ~ exclusion_reason_primary,
    sampleID == 'prominent' & article_id == '29' ~ "RETRACTED", # the article was withdrawn, retracted seems like the closest fit among out labels
    sampleID == 'prominent' & article_id == '37' ~ exclusion_reason_primary,
    sampleID == 'prominent' & article_id == '39' ~ exclusion_reason_primary,
    sampleID == 'prominent' & article_id == '46' ~ exclusion_reason_primary,
    sampleID == 'prominent' & article_id == '54' ~ exclusion_reason_secondary,
    TRUE ~ exclusion_reason_final # if none of the above, retain whatever is in this column already
  ))
```

@@

```{r}
d_articles %>% filter(exclusion_reason_final == "OTHER") %>% select(article_id, sampleID)
```

@@

```{r}
d_articles <- d_articles %>%
  mutate(exclusion_reason_final = case_when(
    sampleID == 'random' & article_id == '68' ~ "NO ACCESS",
    sampleID == 'prominent' & article_id == '2' ~ "CONFERENCE ABSTRACT",
    sampleID == 'prominent' & article_id == '56' ~ "CONFERENCE ABSTRACT",
    sampleID == 'prominent' & article_id == '85' ~ "CONFERENCE ABSTRACT",
    TRUE ~ exclusion_reason_final # if none of the above, retain whatever is in this column already
  ))
```

@@

```{r}
d_articles <- d_articles %>%
  mutate(ppc_linked_final = case_when(
    exclusion_final == "EXCLUDE" ~ NA_character_, # if article is excluded, use NA
    sampleID == 'prominent' & article_id == '38' ~ ppc_linked_secondary,
    exclusion_final == "RETAIN" & exclusion_primary == "RETAIN" ~ ppc_linked_primary,
    exclusion_final == "RETAIN" & exclusion_secondary == "RETAIN" ~ ppc_linked_secondary,
    TRUE ~ "ERROR" # if none of the above, retain whatever is in this column already
  ))
```

@@

```{r}
d_journals_final_eligible <- d_journals_final_eligible %>%
  pivot_longer( # pivot to long format
    cols = c(-journal,-sampleID, -anyPPC,-general_note),
    names_to = c("PPC_type", ".value"), 
    names_sep = "_ppc_"
  )
```

@@

# Journals data — add JIF and COPE

There is some information about the sample that isn't included in the extracted data files (e.g., JIFs, Web of Science subject category), so we need to load some of that information from the original sample files.

```{r}
# list of journals in random sample
journal_list_random <- read_csv(here('data','prepareSample','journals', '03 - final', 'journals-random.csv'), show_col_types = F) %>%
  slice_head(n = 108) %>%
  select(`Journal`)

# list of journals in prominent sample
journal_list_prominent <- read_csv(here('data','prepareSample','journals', '03 - final', 'journals-prominent.csv'), show_col_types = F) %>%
  slice_head(n = 114) %>%
  select('Journal' = `Journal name`)

# list of all prominent journals (some not in sample)
journal_list_prominent_all <- read_csv(here('data','prepareSample','journals', '03 - final', 'journals-prominent.csv'), show_col_types = F)

# list of all random journals (some not in sample)
journal_list_random_all <- read_csv(here('data','prepareSample','journals', '03 - final', 'journals-random.csv'), show_col_types = F)
```

## Identify JIFs for journals

For the random journals we can find the JIFs in the prominent journal sample. Some journals do not have a JIF because they haven't been indexed long enough. We manually checked all of these journals. One JIF needed to be added manually for the journal "Journal of Psychology and Theology". The 2021 JIF for this journal is .80.

```{r}
d_journals_prominent <- d_journals %>%
  filter(sampleID %in% c('prominent', 'both'))

d_journals_random <- d_journals %>%
  filter(sampleID == 'random')

journal_list_prominent_all <- journal_list_prominent_all %>% 
  select(journal = `Journal name`, `2021 JIF`, WOS_first_psych_category)

d_journals_prominent <- left_join(d_journals_prominent, journal_list_prominent_all, by = 'journal')

journal_list_random_all <- journal_list_random_all %>%
  select(journal = Journal, WOS_first_psych_category)

d_journals_random <- left_join(d_journals_random, journal_list_random_all, by = 'journal')

journal_list_prominent_all <- journal_list_prominent_all %>%
  select(-WOS_first_psych_category)

d_journals_random <- left_join(d_journals_random, journal_list_prominent_all, by = 'journal')

tmp <- d_journals_random %>% filter(is.na(`2021 JIF`)) %>% distinct(journal, .keep_all = T)
write_csv(tmp,here('data', 'prepareSample', 'journals', '02 - modified', 'we_need_jifs.csv'))
```

#Add COPE status FROM PRACTICE CODING, does not alter main d_journals tibble
```{r}
# Load googlesheet data

dj_prominent <- read_csv(here('data','raw', 'journals-prominent.csv')) %>% #loads csv of googlesheet data of prominent journals
  slice_head(n = 114) %>%
  select('journal' = 'journal_name', 'COPE' = 'COPE (T = member; F = not a member)' )

dj_random <- read_csv(here('data','raw', 'journals-random.csv')) %>% #loads csv of googlesheet data of random journals
  slice_head(n = 108) %>%
  select('journal' = 'journal_name', 'COPE' = 'COPE (T = member; F = not a member)' )

# Combine above CSVs to one tibble

dj_cope <- bind_rows(dj_prominent, dj_random)

# Identify duplicate rows
duplicates <- duplicated(dj_cope)

# Select the duplicate rows
duplicate_rows <- dj_cope[duplicates, ]

# Check if duplicate rows have the same COPE values for journals
for (i in seq_len(nrow(duplicate_rows))) {
  row <- duplicate_rows[i, ]
  id <- row$ID
  value <- row$Value
  same_values <- all(duplicate_rows$Value[duplicate_rows$ID == id] == value)

  if (same_values) {
    cat("Row", i, "with ID =", id, "has the same values in both columns.\n")
  } else {
    cat("Row", i, "with ID =", id, "has different values in both columns.\n")
  }
}

#All 14 duplicated (both sample) journals had consistent COPE status, remove duplicates
dj_cope <- distinct(dj_cope, journal, .keep_all = T)

#join with main data tibble:
d_primaryONLY <- left_join(d_primaryONLY, dj_cope, by = 'journal')

>>>>

```{r}
d_journals <- d_journals %>%
  select(timestamps = Timestamp,
         sample_id,
         coder_id = `Coder's Initials`,
         journal = `Journal Name`,
         journal_empirical = `Does the journal publish empirical research?`,
         anyPPC = `Check the article types page and check for web comments — does the journal offer any type of PPC?`,
         A_ppc_name = `A. Enter the name of the PPC type offered by the journal:`,
         B_ppc_name = `B. Enter the name of the PPC type offered by the journal:`,
         C_ppc_name = `C. Enter the name of the PPC type offered by the journal:`,
         A_ppc_description = `A. Enter the verbatim description of this type of PPC provided by the journal`,
         B_ppc_description = `B. Enter the verbatim description of this type of PPC provided by the journal`,
         C_ppc_description = `C. Enter the verbatim description of this type of PPC provided by the journal`,
         A_ppc_length = `A. Are there any length limits for this type of PPC?`,
         B_ppc_length = `B. Are there any length limits for this type of PPC?`,
         C_ppc_length = `C. Are there any length limits for this type of PPC?`,
         A_ppc_time = `A. Are there any time limits for submission of this type of PPC?`,
         B_ppc_time = `B. Are there any time limits for submission of this type of PPC?`,
         C_ppc_time = `C. Are there any time limits for submission of this type of PPC?`,
         A_ppc_ref = `A. Are there any reference limits for this type of PPC?`,
         B_ppc_ref = `B. Are there any reference limits for this type of PPC?`,
         C_ppc_ref = `C. Are there any reference limits for this type of PPC?`,
         A_ppc_review = `A. Is this type of PPC sent for independent external peer review?`,
         B_ppc_review = `B. Is this type of PPC sent for independent external peer review?`,
         C_ppc_review = `C. Is this type of PPC sent for independent external peer review?`,
         A_ppc_note = `A. Anything additional, unusual, or interesting to note?`,
         B_ppc_note = `B. Anything additional, unusual, or interesting to note?`,
         C_ppc_note = `C. Anything additional, unusual, or interesting to note?`,
         general_note = `Anything additional, unusual, or interesting to note?`,
         -starts_with("D. "),
         -ends_with("Are there any other types of PPC in this journal? (Remember to check for web comments!)`"))

d_articles <- d_articles %>%
  select(
    sample_id,
    article_id = `article id`,
    coder_id_primary = `primary coder initials`,
    coder_id_secondary = `secondary coder initials`,
    exclusion_primary = `1_exclusion`,
    exclusion_reason_primary = `1_exclusionReason`,
    ppc_linked_primary = `1_Is article linked to PPC`,
    notes_primary = `1_Additional notes`,
    exclusion_secondary = `2_exclusion`,
    exclusion_reason_secondary = `2_exclusionReason`,
    ppc_linked_secondary = `2_Is article linked to PPC`,
    notes_secondary = `2_Additional notes`
  )
```

@@

@@

# Articles data — Resolve coding disagreements

@@

```{r}
# create exclusion_final column
d_articles <- d_articles %>%
  mutate(exclusion_final = case_when(
    sample_id == 'random' & article_id == '37' ~ exclusion_primary,
    sample_id == 'random' & article_id == '41' ~ exclusion_primary,
    sample_id == 'random' & article_id == '68' ~ exclusion_secondary,
    sample_id == 'random' & article_id == '95' ~ exclusion_secondary,
    sample_id == 'random' & article_id == '103' ~ exclusion_primary,
    sample_id == 'random' & article_id == '120' ~ exclusion_primary,
    sample_id == 'prominent' & article_id == '30' ~ exclusion_secondary,
    sample_id == 'prominent' & article_id == '49' ~ exclusion_primary,
    sample_id == 'prominent' & article_id == '66' ~ exclusion_primary,
    sample_id == 'prominent' & article_id == '98' ~ exclusion_primary,
    TRUE ~ exclusion_primary
  ))

# create exclusion_reason_final column
d_articles <- d_articles %>%
  mutate(exclusion_reason_final = case_when(
    sample_id == 'random' & article_id == '37' ~ exclusion_reason_primary,
    sample_id == 'random' & article_id == '41' ~ exclusion_reason_primary,
    sample_id == 'random' & article_id == '68' ~ exclusion_reason_secondary,
    sample_id == 'random' & article_id == '95' ~ exclusion_reason_secondary,
    sample_id == 'random' & article_id == '103' ~ exclusion_reason_primary,
    sample_id == 'random' & article_id == '120' ~ exclusion_reason_primary,
    sample_id == 'prominent' & article_id == '30' ~ exclusion_reason_secondary,
    sample_id == 'prominent' & article_id == '49' ~ exclusion_reason_primary,
    sample_id == 'prominent' & article_id == '66' ~ exclusion_reason_primary,
    sample_id == 'prominent' & article_id == '98' ~ exclusion_reason_primary,
    TRUE ~ exclusion_reason_primary
  ))
```

@@

```{r}
# we've already dealt with cases where the decision to exclude differs, so we're not looking only at cases where the decision to exclude matches, but the reason for exclusion differs
d_articles %>%
  filter(exclusion_primary == exclusion_secondary, 
         exclusion_reason_primary != exclusion_reason_secondary) %>%
  select(article_id, sample_id, exclusion_reason_primary, exclusion_reason_secondary)
```

We will resolve these by applying the exclusion criteria hierarchically. In other words, we use the highest level of exclusion criteria assigned by one of the two coders.
1. Access
2. English
3. Retracted 
4. Non-empirical

```{r}
# update exclusion_reason_final column when primary and secondary coder disagree on reason to exclude.
d_articles <- d_articles %>%
  mutate(exclusion_reason_final = case_when(
    sample_id == 'random' & article_id == '14' ~ exclusion_reason_primary,
    sample_id == 'random' & article_id == '35' ~ exclusion_reason_primary,
    sample_id == 'random' & article_id == '54' ~ exclusion_reason_secondary,
    sample_id == 'random' & article_id == '73' ~ exclusion_reason_primary,
    sample_id == 'random' & article_id == '78' ~ exclusion_reason_primary,
    sample_id == 'prominent' & article_id == '29' ~ "RETRACTED", # the article was withdrawn, retracted seems like the closest fit among out labels
    sample_id == 'prominent' & article_id == '37' ~ exclusion_reason_primary,
    sample_id == 'prominent' & article_id == '39' ~ exclusion_reason_primary,
    sample_id == 'prominent' & article_id == '46' ~ exclusion_reason_primary,
    sample_id == 'prominent' & article_id == '54' ~ exclusion_reason_secondary,
    TRUE ~ exclusion_reason_final # if none of the above, retain whatever is in this column already
  ))
```

@@

```{r}
d_articles %>% filter(exclusion_reason_final == "OTHER") %>% select(article_id, sample_id)
```

@@

```{r}
d_articles <- d_articles %>%
  mutate(exclusion_reason_final = case_when(
    sample_id == 'random' & article_id == '68' ~ "NO ACCESS",
    sample_id == 'prominent' & article_id == '2' ~ "CONFERENCE ABSTRACT",
    sample_id == 'prominent' & article_id == '56' ~ "CONFERENCE ABSTRACT",
    sample_id == 'prominent' & article_id == '85' ~ "CONFERENCE ABSTRACT",
    TRUE ~ exclusion_reason_final # if none of the above, retain whatever is in this column already
  ))
```

@@

```{r}
d_articles <- d_articles %>%
  mutate(ppc_linked_final = case_when(
    exclusion_final == "EXCLUDE" ~ NA_character_, # if article is excluded, use NA
    sample_id == 'prominent' & article_id == '38' ~ ppc_linked_secondary,
    exclusion_final == "RETAIN" & exclusion_primary == "RETAIN" ~ ppc_linked_primary,
    exclusion_final == "RETAIN" & exclusion_secondary == "RETAIN" ~ ppc_linked_secondary,
    TRUE ~ "ERROR" # if none of the above, retain whatever is in this column already
  ))
```

@@

```{r}
d_journals_final_eligible <- d_journals_final_eligible %>%
  pivot_longer( # pivot to long format
    cols = c(-journal,-sample_id, -anyPPC,-general_note),
    names_to = c("PPC_type", ".value"), 
    names_sep = "_ppc_"
  )
```

@@
